Estimating Fixed Frame Galaxy Magnitudes in the SDSS$^1$

### 2002-04-18 docs/paper/paper.tex
chjamnges
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@104 75dff457-8751-0410-ab1a-cd701654ce6b

\title{Estimating Fixed-frame Galaxy Magnitudes in the SDSS$^1$}

@@

%% Mark off your abstract in the ``abstract'' environment. In the manuscript
%% style, abstract will output a Received/Accepted line after the
%% title and affiliation information. No date will appear since the author
%% does not have this information. The dates will be filled in by the
%% editorial office after submission.
\begin{abstract}
Broad-band measurements of flux for galaxies at different redshifts
measure different regions of the rest-frame galaxy spectrum. Certain
astronomical questions, such as the evolution of the luminosity
function of galaxies, require transforming these inherently
redshift-dependent magnitudes into redshift-independent quantities. To
prepare to address these astronomical questions, investigated in
detail in subsequent papers, we implement here the method of
\citet{csabai00a} for fitting spectral energy distributions (SEDs) to
broad-band photometric observations, in the context of the optical
observations of the Sloan Digital Sky Survey (SDSS). Linear
combinations of four spectral templates can reproduce the five SDSS
magnitudes of all galaxies to the precision of the
photometry. Expressed in the appropriate coordinate system, the
distribution of the coefficients multiplying the templates is planar,
and in fact nearly linear. The resulting reconstructed SEDs can be
used to recover fixed-frame magnitudes over a range of redshifts. This
process yields consistent results, in the sense that within each
sample the intrinsic colors of similar type galaxies are nearly
constant with redshift. We compare our results to simpler
interpolation methods and galaxy spectrophotometry.  The software that
generates these results is publicly available and easily adapted to
handle a wide range of galaxy observations.
\end{abstract}

@@

In the future, observations of galaxies (and indeed of any
astronomical sources) will be performed using integral field
spectrographs or equivalent devices which combine high spatial
resolution and high spectral resolution. By that time, the phenomena
of broad-band filters and the quaint terminology surrounding their
usage --- magnitudes, $K$-corrections, color terms, {\it etc.} ---
will have long since been forgotten. However, today, virtually all
observations of galaxies are made through such filters, and special
care has to be taken to recover knowledge of galaxy spectral energy
distributions (SEDs) from these observations.  Reconstructing galaxy
SEDs is nontrivial because SEDs of galaxies contain important
information on scales considerably smaller than the width of typical
broad-band filters. Furthermore, SEDs of distant galaxies are
redshifted such that the more distant the galaxy, the further the
observed bandpass is blueshifted relative to the rest-frame spectral
energy distribution of the object observed.  This paper focuses on a
solution to these problems which (in the optical wavelength regime) is
fast, robust, and consistent over a large range of redshifts.

In the past, people have accounted for these effects using
``$K$-corrections'' applied to the observed magnitudes. In these
analyses, a function $K(z)$ is added to the standard cosmological
bolometric distance modulus $\mathrm{DM}(z)$. Sometimes a single
function has been applied regardless of the galaxy type, though more
recently it has become standard to use a discrete set of $K(z)$
functions depending on galaxy type (based either on morphology or
spectral features). As described in \cite{oke68a} and later papers,
the function is based on the projection of an assumed galaxy SED
redshifted to $z$ onto the measured instrument response as a function
of wavelength (including the effects of the atmosphere, the
reflectivity of the mirrors, the filter transmission, and the response
of the CCD device or photographic emulsion). For example, for the
Sloan Digital Sky Survey (SDSS) filter system, \citet{frei94a} and
\citet{fukugita95a} have both presented results for the
$K$-corrections of galaxies. Each of these papers tabulates $K(z)$
(called $k(z)$ by \citealt{frei94a}) for galaxies of various
morphological types, based on galaxy spectrophotometry.\footnote{Note
that the description on p. 57 of \citet{binney98a} of the meaning of
$k(z)$ in \citet{frei94a} is incorrect; in fact, $k(z)\equiv K(z)$.}

However, galaxies do not all have the same SED, nor are they selected
from some discrete set of SEDs. For this reason, these schemes for
applying the $K$-corrections can fail to be self-consistent: the SED
assumed for the $K$-corrections can be significantly inconsistent with
the observed galaxy colors! As we demand more precision from our
astronomical data analysis in new, large multi-band surveys, and in
particular as we try to quantify the evolution of galaxies, we must
take a more sophisticated approach to approximating fixed-frame
observations of galaxies.

Our approach here is to use a method for inferring the underlying SEDs
of a set of galaxies at a range of redshifts by requiring that their
SEDs all be drawn from a similar population. For each galaxy, we will
recover a model SED, which can be used to synthesize the galaxy's
magnitude in any bandpass. Although the operation we are performing on
the magnitudes is not a ``$K$-correction'' in the historical sense of
the term, we will refer to it as such in this and subsequent
papers. Our approach is equivalent (nearly identical) to the
photometric metric redshift estimation methods of \citet{csabai00a},
except we use slightly different coordinate systems for our SED
template space. In fact, the software includes a fast and relatively
accurate photometric redshift estimator based on their method.

We implemented this system and are publishing it in order to lay the
groundwork for upcoming papers which will rely heavily on the
reliability of the fixed-frame magnitudes determined here.  Because
the SDSS is one of the largest astronomical surveys to date, our
method will be useful to a large number of other investigators, both
inside and outside the SDSS collaboration.  Thus, this paper also
serves the purpose of describing the release of a piece of
software. The computer software we distribute builds itself into a C
shared object library, around which we have written both stand-alone C
programs and IDL routines. We have made the source code available
publicly, through a web page and through a public CVS
repository. Improvements or ports to other languages implemented by
users may be incorporated into the code upon request. The conditions
of use for the code are that this paper is cited in any resulting
refereed journal article and that the version of the code used is
specified in any such paper. The version of the code used to make the
figures for this paper is {\tt kcorrect \kversion}.

We refer throughout this paper to $AB$ magnitudes, meaning that the
magnitudes are designed such that ():
\begin{eqnarray}
\label{ABdef}
m_{AB} &=& -2.41 - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\lambda \lambda f(\lambda) R(\lambda)}
{\int_{0}^{\infty} d\lambda \lambda^{-1} R(\lambda)}\right]
\cr
&=& -48.60 - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\nu \nu^{-1} f(\nu) R(\nu)}
{\int_{0}^{\infty} d\nu \nu^{-1} R(\nu)}\right],
\end{eqnarray}
where $R(\lambda)$ is the fraction of photons entering the Earth's
atmosphere which are detected as a function of wavelength (a unitless
quantity). In this equation, $f(\lambda)$ is in units of ergs
cm$^{-2}$ s$^{-1}$ \AA$^{-1}$, and $f(\nu)$ is in units of ergs
cm$^{-2}$ s$^{-1}$ Hz$^{-1}$. The normalizations defined here mean
that an object with $f(\nu) = 3631\mathrm{~Jy} = 3.631 \times
10^{-20}$ ergs cm$^{-2}$ s$^{-1}$ Hz$^{-1}$ has all its AB magnitudes
equal to zero.  The difference in the zeropoint in the two equations
simply corresponds to the factor of the speed of light $c$ (expressed
in \AA\ s$^{-1}$) between the frequency and wavelength scales.

Section \ref{sedfit} describes our method of fitting galaxy SEDs to
broad-band photometry and how to calculate $K$-corrections. Section
\ref{data} applies the method to galaxies in the SDSS, showing that
the fits are robust.  Section \ref{conclusions} concludes and
discusses future development of the method described here.

\section{Calculating Fixed-frame Galaxy Magnitudes}
\label{sedfit}

First, we describe how we reconstruct galaxy SEDs from broad-band
magnitude measurements. Second, we describe how we convert these SEDs
into estimates of fixed-frame galaxy magnitudes.

@@

Our task is to recover a model for the galaxy SED from broad-band
photometric measurements. Since a set of broad band galaxy fluxes does
not correspond uniquely to a particular SED, and because galaxy SEDs
are known to have significant structure over wavelength ranges small
compared to our bandpass, this task is an ill-posed, inverse problem:
given any set of observed galaxy colors, there is not a unique SED to
which those colors correspond, even in the absence of errors. However,
we are not completely ignorant about the forms which galaxy SEDs take,
so we can attempt to use what we know about galaxy SEDs to simplify
our task.  The method described here for doing so follows closely the
methods of estimating photometric redshifts used by \citet{csabai00a}
and \citet{budavari00a}. It is designed to take advantage of what we
already understand about galaxy spectral energy distributions.
%In addition, as implemented here, the method is
%equivalent to a fit to the star-formation history of the galaxies
%using a spectrophotometric model; we will investigate the implications
%of the model fits in a separate paper and concentrate here on the
%empirical aspects of the results.

Begin with an SED space defined by $N_b$ template galaxy SEDs
$\vv{v}_i(\lambda)$ (say from \citealt{bruzual93a}), where $N_b$ is
large. In principle, this should be a complete set of galaxy SEDs if
we want {\it all} galaxy SEDs to be in the space spanned by the
$\vv{v}_i(\lambda)$; in practice, this property is not necessary.
Given a set of $\vv{v}_i(\lambda)$, one can find an orthonormal set of
spectra $\vv{b}_i(\lambda)$ spanning the same space.  To be more
precise, we define the dot-product in the space of galaxy spectra to
be:
\begin{equation}
\vv{v}_i \cdot \vv{v}_j =
\int_{\lambda_{\mathrm{min}}}^{\lambda_{\mathrm{max}}} d\lambda
v_i(\lambda) v_j(\lambda),
\end{equation}
where $\lambda_{\mathrm{min}}$ and $\lambda_{\mathrm{max}}$ define the
wavelength range used to define orthogonality. These limits are
defined not over the full range covered by the $\vv{v}_i(\lambda)$
spectra, but instead over a smaller range which is within observable
wavelengths over most of the redshift range of the
sample. \footnote{Were we to define orthogonality over the full range
of $\vv{v}(\lambda)$, much of which is unobservable, the geometry of
the distribution of galaxies in this space would be more complicated.
If the template spectra were identical in the subrange chosen and only
differed outside of it, our choice would cause computational
problems, but in fact the spectra are all independent in the
restricted, as well as the full, wavelength range.}  We then define
the basis of the SED space such that the $N_b$ template SEDs,
$\vv{v}_i$, are linear combinations of $N_b$ basis spectra $\vv{b}_i$
and that $\vv{b}_i \cdot \vv{b}_j = \delta_{ij}$. These conditions do
not fully specify the $\vv{b}_i$, which naturally may be rotated
arbitrarily within our $N_b$-dimensional subspace of spectral space.

In our case the number of observed bands is less than $N_b$, so we
cannot determine an individual galaxy's position in this
$N_b$-dimensional space from its broad-band colors alone.  However, as
outlined by \citet{csabai00a}, one can find a lower dimensional space
defined by $N_t$ vectors $\vv{e}_k$ which best fits the {\it set} of
galaxies, as follows. The SED of galaxy $i$ may be reconstructed from a
linear combination of the basis spectra in this lower dimensional
space:
\begin{equation}
\label{lincomb}
f_{\mathrm{rec},i}(\lambda) d\lambda = 
\sum_k a_{i,k} \sum_j e_{k,j} b_j(\lambda) d\lambda,
\end{equation}
where $a_k$ are the components of the galaxy projected in the
low-dimensional space. From this SED it is easy to determine the
reconstructed flux
$F_{\mathrm{rec},il}$ in each survey bandpass by projecting
the model SED onto the bandpass $l$: 
\begin{equation}
F_{\mathrm{rec},il} = \int_0^\infty d\lambda
f_{\mathrm{rec},i}(\lambda) S_l(\lambda),
\end{equation}
where $S_l(\lambda)$ is the response of the instrument in band $l$.  We define
\begin{equation}
\chi^2 = \sum_i \sum_l
\frac{(F_{\mathrm{obs},il}-F_{\mathrm{rec},il})^2}{\sigma_{il}^2},
\end{equation}
where $\sigma_{il}$ is the estimated error in the observed flux
$F_{\mathrm{obs},il}$ in bandpass $l$ for galaxy $i$. Taking
derivatives of $\chi^2$ with respect to $a_{i,k}$ and $e_{k,j}$
results in a set of equations bilinear in $a_{i,k}$ and $e_{k,j}$.
\citet{csabai00a} describe how to iteratively solve this bilinear
equation by first fixing $e_{k,j}$ and fitting for $a_{i,k}$, then
fixing $a_{i,k}$ and fitting for $e_{k,j}$, and iterating that
procedure.

@@

We define the ``total flux'' $f_k$ of each template as the flux in the
range $\lambda_{\mathrm{min}}<\lambda<\lambda_{\mathrm{max}}$; then an
estimate of the flux in this range is $F_{{t}}=\Sigma_k a_k f_k$.
Galaxies of some fixed $F_t$ clearly lie on a plane in the component
space $a_k$. Therefore, a natural choice for the orientation of the
axes $\vv{e}_k$ is to let $\vv{e}_0$ be perpendicular to the planes of
constant $F_t$; in this manner, the coefficient $a_0$ in Equation
(\ref{lincomb}) is directly proportional to $F_t$ and is thus purely a
measure of the object's flux (in the wavelength range
$\lambda_{\mathrm{min}} <\lambda < \lambda_{\mathrm{max}}$), while the
parameters $a_i/a_0$ (where $i>0$) primarily measure the shape of the
galaxy SED in the wavelength range $\lambda_{\mathrm{min}} <\lambda <
\lambda_{\mathrm{max}}$. $F_t$ can be trivially related to $L_t$, the
total luminosity in this range, by the inverse square of the
luminosity distance (from, for example, \citealt{hogg99a}).  As it
turns out, the parameters $a_i/a_0$ tend to be distributed
approximately in an ellipsoid, with not much curvature within the
space defined by the template SEDs. Thus, we transform the axes such
that $a_i/a0 = 0$ is the mean of the distribution and rotate the axes
$\vv{e}_i$ (for $i>0$) such that they are aligned with the principal
axes of the ellipsoid. Furthermore, we normalize the vectors
$\vv{e}_i$ such that the zeroth component $a_0=F_t$ is output by the
code in units of ergs cm$^{-2}$ s$^{-1}$ (within the range
$\lambda_{\mathrm{min}} <\lambda < \lambda_{\mathrm{max}}$).

In this way, we can reconstruct SEDs based on the broad-band
magnitudes of galaxies. From these reconstructed SEDs one can estimate
$K$-corrections, develop a measure of galaxy spectral type, or
synthesize other measurements of galaxy flux.  The templates
determined during the procedure can, of course, be used to estimate
photometric redshifts, as \citet{csabai00a} describes.

\subsection{Applying Direct Constraints on $a_i/a0$}
\label{direct}

Sometimes we will want to apply direct constraints on the coefficients
$a_i/a_0$ of the galaxy we are fitting. For example, in this paper we
fit our templates to a set of galaxies with high quality photometry in
all bands. However, we would like to use these templates to fit for
the SEDs of other galaxies, which might have much worse photometry in
one or more bands. Since the errors in the photometry can be
catastrophic (that is, non-gaussian), when we fit the SEDs of the
lower signal-to-noise galaxies it makes sense to at least weakly
constrain these galaxies to have reasonable SEDs, in the sense that
the $a_i/a_0$ ought not to be too different than for well-measured
galaxies. Here we describe how to implement direct constraints on the
coefficients $a_i/a_0$ which describe the galaxy SED.

We will assume that any constraints can be expressed as a combination
of gaussian constraints with means $a_{c,j}/a_{c,0}$, variance matrices
$V_{c,jl}$, weighted by amplitudes $A_c$.  We calculate a
preliminary value of $a_{c,0}$, assuming $a_i = 0$. Then we constrain
our fit by adding a term to $\chi^2$ for each constraint:
\begin{equation}
\sum_j \sum_l V^{-1}_{jl} (a_l/a_{c,0}-a_{c,l}/a_{c,0})
(a_j/a_{c,0}-a_{c,j}/a_{c,0})  
\end{equation}
The resulting $\chi^2$ can still be minimized linearly, so we have
sacrificed very little in speed.

Note that by using this constraint, one can calculate $K$-corrections
even for galaxies with only one or two bands measured, and still
obtain sensible results.

@@

Given AB galaxy magnitudes $^{z}m_{AB}$ and errors for a galaxy at
redshift $z$, one can use the method of the previous subsection to
reconstruct the galaxy rest-frame SED.  With this reconstruction in
hand, it is straightforward to project the SED onto any bands at any
redshift $z_0$ ({\it cf.} Equation \ref{ABdef}):
\begin{eqnarray}
\band{z_0}{m_{\mathrm{rec}}} &=& -2.41 - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\lambda \frac{\lambda}{1+z_0}
f_{\mathrm{rec}}(\frac{\lambda}{1+z_0}) R(\lambda)} {\int_{0}^{\infty}
d\lambda \lambda^{-1} R(\lambda)}\right] \cr &=& -48.60 - 2.5
\log_{10}\left[ \frac{\int_{0}^{\infty} d\nu \nu^{-1} (1+z_0)
f_{\mathrm{rec}}[\nu (1+z_0)] R(\nu)} {\int_{0}^{\infty} d\nu \nu^{-1}
R(\nu)}\right]
\end{eqnarray}
The simple way to remember where the factors of $(1+z)$ appear in
these equations is to remember that $K$-corrections are defined to be
the change in magnitude with redshift for {\it fixed bolometric flux}.
So when the spectrum is stretched by the $(1+z_0)$ redshift factor, as
in $f(\lambda/(1+z_0))$, the bolometric flux should be constant; that
is, you must apply the extra multiplicative factor $1/(1+z_0)$ to
counteract the stretching of the spectrum.

>>>>

\title{Estimating Fixed Frame Galaxy Magnitudes in the SDSS$^1$}

@@

%% Mark off your abstract in the ``abstract'' environment. In the manuscript
%% style, abstract will output a Received/Accepted line after the
%% title and affiliation information. No date will appear since the author
%% does not have this information. The dates will be filled in by the
%% editorial office after submission.
\begin{abstract}
Broad-band measurements of flux for galaxies at different redshifts
measure different regions of the rest-frame galaxy spectrum. Certain
astronomical questions, such as the evolution of the luminosity
function of galaxies, require transforming these inherently
redshift-dependent magnitudes into redshift-independent quantities. To
prepare to address these astronomical questions, investigated in
detail in subsequent papers, we fit spectral energy distributions
(SEDs) to broad band photometric observations, in the context of the
optical observations of the Sloan Digital Sky Survey (SDSS). Linear
combinations of four spectral templates can reproduce the five SDSS
magnitudes of all galaxies to the precision of the
photometry. Expressed in the appropriate coordinate system, the
distribution of the coefficients multiplying the templates is planar,
and in fact nearly linear. The resulting reconstructed SEDs can be
used to recover fixed frame magnitudes over a range of redshifts. This
process yields consistent results, in the sense that within each
sample the intrinsic colors of similar type galaxies are nearly
constant with redshift. We compare our results to simpler
interpolation methods and galaxy spectrophotometry from the SDSS.  The
software that generates these results is publicly available and easily
adapted to handle a wide range of galaxy observations.
\end{abstract}

@@

In the future, observations of galaxies (and indeed of any
astronomical sources) will be performed using integral field
spectrographs or equivalent devices which combine high spatial
resolution and high spectral resolution. By that time, the phenomena
of broad band filters and the quaint terminology surrounding their
usage --- magnitudes, $K$-corrections, color terms, {\it etc.} ---
will have long since been forgotten. However, today, virtually all
observations of galaxies are made through such filters, and special
care has to be taken to recover knowledge of galaxy spectral energy
distributions (SEDs) from these observations.  Reconstructing galaxy
SEDs is nontrivial because SEDs of galaxies contain important
information on scales considerably smaller than the width of typical
broad band filters. Furthermore, SEDs of distant galaxies are
redshifted such that the more distant the galaxy, the further the
observed bandpass is blueshifted relative to the rest-frame spectral
energy distribution of the object observed.  This paper focuses on a
solution to these problems which (in the optical wavelength regime) is
fast, robust, and consistent over a large range of redshifts.

In the past, people have accounted for these effects using
``$K$-corrections'' applied to the observed magnitudes. In these
analyses, a function $K(z)$ is added to the standard cosmological
bolometric distance modulus $\mathrm{DM}(z)$ to obtain the
relationship between the apparent magnitude $m$ and the absolute
magnitude $M$: 
\begin{equation}
m = M + \mathrm{DM}(z) + K(z)
\end{equation}
Sometimes a single function has been applied regardless of the galaxy
type, though more recently it has become standard to use a discrete
set of $K(z)$ functions depending on galaxy type (based either on
morphology or spectral features). As described in \cite{oke68a} and
later papers, the function is based on the projection of an assumed
galaxy SED redshifted to $z$ onto the measured instrument response as
a function of wavelength (including the effects of the atmosphere, the
reflectivity of the mirrors, the filter transmission, and the response
of the CCD device or photographic emulsion). For example, for the
Sloan Digital Sky Survey (SDSS) filter system, \citet{frei94a} and
\citet{fukugita95a} have both presented results for the
$K$-corrections of galaxies. Each of these papers tabulates $K(z)$
(called $k(z)$ by \citealt{frei94a}) for galaxies of various
morphological types, based on galaxy spectrophotometry.\footnote{Note
that the description on p. 57 of \citet{binney98a} of the meaning of
$k(z)$ in \citet{frei94a} is incorrect; in fact, their $k(z)$ is
exactly equivalent to our $K(z)$.}.

However, galaxies do not all have the same SED, nor are they selected
from some discrete set of SEDs. For this reason, these schemes for
applying the $K$-corrections can fail to be self-consistent: the SED
assumed for the $K$-corrections can be significantly inconsistent with
the observed galaxy colors! As we demand more precision from our
astronomical data analysis in new, large multi-band surveys, and in
particular as we try to quantify the evolution of galaxies, we must
take a more sophisticated approach to approximating fixed frame
observations of galaxies.

Our approach here is to use a method for inferring the underlying SEDs
of a set of galaxies at a range of redshifts by requiring that their
SEDs all be drawn from a similar population. For each galaxy, we will
recover a model SED, which can be used to synthesize the galaxy's
magnitude in any bandpass. Although the operation we are performing on
the magnitudes is not a ``$K$-correction'' in the historical sense of
the term, we will refer to it as such in this and subsequent
papers. Our approach is equivalent (nearly identical) to the
photometric redshift estimation methods of \citet{csabai00a},
except we use slightly different coordinate systems for our SED
template space. In fact, the software includes a fast and relatively
accurate photometric redshift estimator based on their method.

We implemented this system and are publishing it in order to lay the
groundwork for upcoming papers which will rely heavily on the
reliability of the fixed frame magnitudes determined here.  Because
the SDSS is one of the largest astronomical surveys to date, our
method will be useful to a large number of other investigators, both
inside and outside the SDSS collaboration.  Thus, this paper also
serves the purpose of describing the release of a piece of
software. The computer software we distribute builds itself into a C
shared object library, around which we have written both stand-alone C
programs and IDL routines. We have made the source code available
publicly, through a web page and through a public CVS
repository. Improvements or ports to other languages implemented by
users may be incorporated into the code upon request. The conditions
of use for the code are that this paper is cited in any resulting
refereed journal article and that the version of the code used is
specified in any such paper. The version of the code used to make the
figures for this paper is {\tt kcorrect \kversion}.

We refer throughout this paper to $AB$ magnitudes, meaning that the
magnitudes are designed such that (\citealt{fukugita96a}):
\begin{eqnarray}
\label{ABdef}
m_{AB} &=& -2.41 - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\lambda \lambda f(\lambda) R(\lambda)}
{\int_{0}^{\infty} d\lambda \lambda^{-1} R(\lambda)}\right]
\cr
&=& -48.60 - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\nu \nu^{-1} f(\nu) R(\nu)}
{\int_{0}^{\infty} d\nu \nu^{-1} R(\nu)}\right],
\end{eqnarray}
where $R(\lambda)$ is the fraction of photons entering the Earth's
atmosphere which are detected as a function of wavelength (a unitless
quantity). In this equation, $f(\lambda)$ is in units of ergs
cm$^{-2}$ s$^{-1}$ \AA$^{-1}$, and $f(\nu)$ is in units of ergs
cm$^{-2}$ s$^{-1}$ Hz$^{-1}$. The normalizations defined here mean
that an object with $f(\nu) = 3631\mathrm{~Jy} = 3.631 \times
10^{-20}$ ergs cm$^{-2}$ s$^{-1}$ Hz$^{-1}$ has all its AB magnitudes
equal to zero.  The difference in the zeropoint in the two equations
simply corresponds to the factor of the speed of light $c$ (expressed
in \AA\ s$^{-1}$) between the frequency and wavelength scales.

Section \ref{sedfit} describes our method of fitting galaxy SEDs to
broad band photometry and how to calculate $K$-corrections. Section
\ref{data} applies the method to galaxies in the SDSS, showing that
the fits are robust.  Section \ref{conclusions} concludes and
discusses future development of the method described here.

\section{Calculating Fixed Frame Galaxy Magnitudes}
\label{sedfit}

First, we describe how we reconstruct galaxy SEDs from broad band
magnitude measurements. Second, we describe how we convert these SEDs
into estimates of fixed frame galaxy magnitudes.

@@

Our task is to recover a model for the galaxy SED from broad band
photometric measurements. Since a set of broad band galaxy fluxes does
not correspond uniquely to a particular SED, and because galaxy SEDs
are known to have significant structure over wavelength ranges small
compared to our bandpass, the inverse problem of reconstructing SEDs
from broad band fluxes is ill-posed.  However, we are not completely
ignorant about the forms which galaxy SEDs take, so we can attempt to
use what we know about galaxy SEDs to simplify our task.  The method
described here for doing so follows closely the method of estimating
photometric redshifts used by \citet{csabai00a}.  It is designed to
take advantage of what we already understand about galaxy spectral
energy distributions.
%In addition, as implemented here, the method is
%equivalent to a fit to the star-formation history of the galaxies
%using a spectrophotometric model; we will investigate the implications
%of the model fits in a separate paper and concentrate here on the
%empirical aspects of the results.

Begin with an SED space defined by $N_b$ template galaxy SEDs
$\vv{v}_i(\lambda)$ (for example from \citealt{bruzual93a}), where $N_b$ is
large. In principle, this should be a complete set of galaxy SEDs if
we want {\it all} galaxy SEDs to be in the space spanned by the
$\vv{v}_i(\lambda)$; in practice, this property is not necessary.
Given a set of $\vv{v}_i(\lambda)$, one can find an orthonormal set of
spectra $\vv{b}_i(\lambda)$ spanning the same space.  To be more
precise, we define the dot-product in the space of galaxy spectra to
be:
\begin{equation}
\vv{v}_i \cdot \vv{v}_j =
\int_{\lambda_{\mathrm{min}}}^{\lambda_{\mathrm{max}}} d\lambda
v_i(\lambda) v_j(\lambda),
\end{equation}
where $\lambda_{\mathrm{min}}$ and $\lambda_{\mathrm{max}}$ define the
wavelength range used to define orthogonality. These limits are
defined not over the full range covered by the $\vv{v}_i(\lambda)$
spectra, but instead over a smaller range which is within observable
wavelengths over most of the redshift range of the sample.\footnote{
If the template spectra were identical in the subrange chosen and only
differed outside of it, our choice would cause computational problems,
but in fact the spectra are all independent in the restricted, as well
as the full, wavelength range.}  We then define the basis of the SED
space such that the $N_b$ template SEDs, $\vv{v}_i$, are linear
combinations of $N_b$ basis spectra $\vv{b}_i$ and that $\vv{b}_i
\cdot \vv{b}_j = \delta_{ij}$. These conditions do not fully specify
the $\vv{b}_i$, which naturally may be rotated arbitrarily within our
$N_b$-dimensional subspace of spectral space.

In our case the number of observed bands is less than $N_b$, so we
cannot determine an individual galaxy's position in this
$N_b$-dimensional space from its broad band colors alone.  However, as
outlined by \citet{csabai00a}, one can find a lower dimensional space
defined by $N_t$ vectors $\vv{e}_k$ which best fits the {\it set} of
galaxies, as follows. The SED of galaxy $i$ may be reconstructed from a
linear combination of the basis spectra in this lower dimensional
space:
\begin{equation}
\label{lincomb}
f_{\mathrm{rec},i}(\lambda) d\lambda = 
\sum_k a_{i,k} \sum_j e_{k,j} b_j(\lambda) d\lambda,
\end{equation}
where $a_k$ are the components of the galaxy projected in the
low-dimensional space. From this SED it is easy to determine the
reconstructed flux
$F_{\mathrm{rec},il}$ in each survey bandpass by projecting
the model SED onto the bandpass $l$: 
\begin{equation}
F_{\mathrm{rec},il} = \int_0^\infty d\lambda
f_{\mathrm{rec},i}(\lambda) S_l(\lambda),
\end{equation}
where $S_l(\lambda)$ is the response of the instrument in band $l$.  We define
\begin{equation}
\chi^2 = \sum_i \sum_l
\frac{(F_{\mathrm{obs},il}-F_{\mathrm{rec},il})^2}{\sigma_{il}^2},
\end{equation}
where $\sigma_{il}$ is the estimated error in the observed flux
$F_{\mathrm{obs},il}$ in bandpass $l$ for galaxy $i$. Taking
derivatives of $\chi^2$ with respect to $a_{i,k}$ and $e_{k,j}$
results in a set of equations bilinear in $a_{i,k}$ and $e_{k,j}$.
\citet{csabai00a} describe how to iteratively solve this bilinear
equation by first fixing $e_{k,j}$ and fitting for $a_{i,k}$, then
fixing $a_{i,k}$ and fitting for $e_{k,j}$, and iterating that
procedure.

@@

We define the ``total flux'' $f_k$ of each template as the flux in the
range $\lambda_{\mathrm{min}}<\lambda<\lambda_{\mathrm{max}}$; then an
estimate of the flux in this range is $F_{{t}}=\Sigma_k a_k f_k$.
Galaxies of some fixed $F_t$ clearly lie on a plane in the component
space $a_k$. Therefore, a natural choice for the orientation of the
axes $\vv{e}_k$ is to let $\vv{e}_0$ be perpendicular to the planes of
constant $F_t$; in this manner, the coefficient $a_0$ in Equation
(\ref{lincomb}) is directly proportional to $F_t$ and is thus purely a
measure of the object's flux (in the wavelength range
$\lambda_{\mathrm{min}} <\lambda < \lambda_{\mathrm{max}}$), while the
parameters $a_i/a_0$ (where $i>0$) primarily measure the shape of the
galaxy SED in the wavelength range $\lambda_{\mathrm{min}} <\lambda <
\lambda_{\mathrm{max}}$. $F_t$ can be trivially related to $L_t$, the
total luminosity in this range, by the inverse square of the
luminosity distance (from, for example, \citealt{hogg99a}).  As it
turns out, the parameters $a_i/a_0$ tend to be distributed
approximately in an ellipsoid, with not much curvature within the
space defined by the template SEDs. Thus, we transform the axes such
that $a_i/a_0 = 0$ is the mean of the distribution and rotate the axes
$\vv{e}_i$ (for $i>0$) such that they are aligned with the principal
axes of the ellipsoid. Furthermore, we normalize the vectors
$\vv{e}_i$ such that the zeroth component $a_0=F_t$ is output by the
code in units of ergs cm$^{-2}$ s$^{-1}$ (within the range
$\lambda_{\mathrm{min}} <\lambda < \lambda_{\mathrm{max}}$).

In this way, we can reconstruct SEDs based on the broad band
magnitudes of galaxies. From these reconstructed SEDs one can estimate
$K$-corrections, develop a measure of galaxy spectral type, or
synthesize other measurements of galaxy flux.  The templates
determined during the procedure can, of course, be used to estimate
photometric redshifts, as \citet{csabai00a} describes.

\subsection{Applying Direct Constraints on $a_i/a_0$}
\label{direct}

Sometimes we will want to apply direct constraints on the coefficients
$a_i/a_0$ of the galaxy we are fitting. For example, in this paper we
fit our templates to a set of galaxies with high quality photometry in
all bands. However, we would like to use these templates to fit for
the SEDs of other galaxies, which might have much worse photometry in
one or more bands. Since the errors in the photometry can be
catastrophic (that is, non-gaussian), when we fit the SEDs of the
lower signal-to-noise galaxies it makes sense to at least weakly
constrain these galaxies to have reasonable SEDs, in the sense that
the $a_i/a_0$ ought not to be too different than those for
well-measured galaxies. We can assume a gaussian prior probability
distribution with a mean $\avg{a_{c,j}/a_{c,0}}$, a variance matrix
$V_{c,jl}$, and a weight $A_c$.  We calculate a preliminary value of
${\hat{a}}_{c,0}$, assuming $a_i = 0$. Then we constrain our fit by
adding a term to $\chi^2$ as follows:
\begin{equation}
\sum_j \sum_l A_c V^{-1}_{jl}
\left(\frac{a_l}{{\hat{a}}_{c,0}}-\Avg{\frac{a_{c,l}}{a_{c,0}}}\right)
\left(\frac{a_j}{{\hat{a}}_{c,0}}-\Avg{\frac{a_{c,j}}{a_{c,0}}}\right)
\end{equation}
The resulting $\chi^2$ can still be minimized linearly, so we have
sacrificed very little in speed.  Note that with such a constraint,
one can calculate $K$-corrections even for galaxies with only one or
two bands measured, and still obtain sensible results.

@@

Given AB galaxy magnitudes $\band{z}{m}$ and errors for a galaxy at
redshift $z$, one can use the method of the previous subsection to
reconstruct the galaxy rest-frame SED.  With this reconstruction in
hand, it is straightforward to project the SED onto any bands at any
redshift $z_0$ ({\it cf.} Equation \ref{ABdef}):
\begin{eqnarray}
\band{z_0}{m_{\mathrm{rec}}} &=& -2.41 - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\lambda \frac{\lambda}{1+z_0}
f_{\mathrm{rec}}(\frac{\lambda}{1+z_0}) R(\lambda)} {\int_{0}^{\infty}
d\lambda \lambda^{-1} R(\lambda)}\right] \cr &=& -48.60 - 2.5
\log_{10}\left[ \frac{\int_{0}^{\infty} d\nu \nu^{-1} (1+z_0)
f_{\mathrm{rec}}[\nu (1+z_0)] R(\nu)} {\int_{0}^{\infty} d\nu \nu^{-1}
R(\nu)}\right]
\end{eqnarray}
The simple way to remember where the factors of $(1+z)$ appear in
these equations is to remember that $K$-corrections are defined to be
the change in magnitude with redshift for {\it fixed bolometric flux}.
So when the spectrum is stretched by the $(1+z_0)$ redshift factor, as
in $f(\lambda/(1+z_0))$, the bolometric flux should be constant; that
is, you must apply the extra multiplicative factor $1/(1+z_0)$ to
counteract the stretching of the spectrum.

Given the reconstructed magnitudes, one can calculate the fixed frame
magnitudes in one of two ways:
\begin{enumerate}
\item Simply use $\band{z_0}{m_{\mathrm{rec}}}$, the reconstructed
magnitudes at redshift $z_0$, as the fixed frame magnitudes.
\item Use the difference
$\band{z_0}{m_{\mathrm{rec}}}-\band{z}{m_{\mathrm{rec}}}$ as an
estimate of the correction to apply to the observed $\band{z}{m}$.
\end{enumerate}
If the reconstructions $\band{z}{m_{\mathrm{rec}}}$ are very similar
to the actual observations $\band{z}{m}$ (as we will show is the case
for our fits), there is very little difference between these two
methods. However, if the reconstructions are poor, there will be large
differences between the methods; in particular, method (1) will yield
a color distribution of galaxies with artificially low
dimensionality. Having estimated $\band{z_0}{m}$, you can then
calculate the luminosity by simply applying to the calculated
magnitude the cosmological distance modulus (that is, the luminosity
distance-squared law, as tabulated by, {\it e.g.},  \citealt{hogg99a}).
